A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-10-12 paper/paper.tex
Starting RL-sentinel

\subsection{Sentinel Hypothesis}
many investigators speculate that this central neural network reflects the brain’s relentless tracking and prediction of environmental events to adaptively optimize the organism's future action (Schacter et al., 2007).

The contents of memory and prospection may be nec- essarily and adaptively tied to things social in nature, po- tentially the most behaviorally relevant stimuli. As such, the evolutionary advancement of autonoetic awareness would correspond to the neural processes of theory of mind. The utilization of prospection is arguably of greatest strategic importance, particularly in the secur- ing of mates, protection, and resources (Suddendorf \&

Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states, 
Schilbach et al., 2008
Clan mentality
DMN does social cognition -> but it exists in rats and monkeys with much more poorly developped social capacities
-> our computational framework can be equally applied to humans
and other primates.

How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance?

Humans have learned to solve ecological problems
by social means.

Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set? -> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action

 The 'sentinel hypothesis' is a major functional account to explain DMN activity that this canonical network acts as a cohesive unit. This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, \& Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, \& Vogeley, 2008). Acting as a sentinel, DMN functions as a broad focus manager that performs continuous environmental tracking for salient and unexpected events (Buckner et al., 2008). This default activity may allow the detection of possible predators and potentially serves an evolutionarily conserved purpose by means of attention allocation (Raichle et al., 2001). The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, \& Stein, 2007). On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, \& Buckner, 2010). A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, \& Yarkoni, 2014; Schacter, Addis, \& Buckner, 2007). However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, \& Bastianello, 2010; McCormick et al., 2014). Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.

 \begin{itemize}
   \item The sentinel account does not provide a formal mechanisms of
   how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explorated and which decision trees are neglected,
    which implies allocation of attention and computational load.
 \end{itemize}

>>>>

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
functional account that was proposed for the DMN \cite{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflects the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation
\cite{}.
-> the most behaviorally relevant stimuli
-> particularly in the secur- ing of mates, protection, and resources
-> tracking for salient and unexpected events

This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-relevant information processing

Environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\cite{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems in a social way
\cite{whiten1988machiavellian}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecific
turned into a decisive evolutionary advantage.

-> contents of memory and everyday dreaming tied to people and relationships to them

Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states, 
Schilbach et al., 2008
Clan mentality

role increasing during self-referential tasks.

 \begin{itemize}
    \item DMN does social cognition -> but it exists in rats
    and monkeys with much more poorly developped social capacities
    -> our computational framework can be equally applied to humans
    and other primates.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explorated and which decision trees are neglected,
    which implies allocation of attention and computational load.
    \item How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance? Yes, if
    environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set? -> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
 \end{itemize}
